---
title: "FSA_CDL_Crosswalk"
author: "Kaitlyn Spangler"
date: "2024-03-20"
output: html_document
---

This exploration will visualize the condensing of FSA crop categories into the CDL categories used in USDA's CDL CropScape dataset - which is a 30m resolution raster dataset classifying all agricultural land use in the coterminous US.

## Load Data

```{r}
options("install.lock"=FALSE)
#install.packages("fauxnaif", dependencies = T)
library(fauxnaif)
library(svglite)
library(tidyverse)

ap_cso <- read.csv("./data/FSA-to-CDL_Crosswalk_1.csv", na.strings=c("","NA","Blank", "blank"))
ap_cso <- ap_cso[-c(1695:1737), ]
# ensure all NAs match 


coarse <- read.csv("./data/FSA-to-CDL_Crosswalk_2.csv", na.strings=c("","NA","Blank", "blank", "Turn Acres"))
coarse <- coarse[-c(1695:1737), ]
#make "Turn Acres" NA in CDL_Crop_Name - TO DO
  
```

## Summarize FSA data

```{r}
library(ggalluvial)
library(ggplot2)

#Annual and Perennials

sum <- ap_cso %>% group_by(Annual_Perennial) %>% mutate(AP_n = n()) 
list <- c(149, 738, 807)
names <- c("NA", "Perennial", "Annual")

ap <- data.frame(list, names)

ap_plot <- ggplot(ap, aes(x = reorder(names, -list), y = list, fill = names)) +
  geom_bar(stat = "identity") +
  xlab("Type of FSA Crop") +
  ylab("Count")+
  theme_minimal()+
  theme(legend.position = "none")

# Commodity, specialty, other 
sum2 <- ap_cso %>% group_by(Commodity_Specialty_Other) %>% mutate(CSO_n = n()) 
list2 <- c(84, 1142, 387)
names2 <- c("Commodity", "Specialty", "Other")

ap2 <- data.frame(list2, names2)

ap_plot2 <- ggplot(ap2, aes(x = reorder(names2, -list2), y = list2, fill = names2)) +
  geom_bar(stat = "identity") +
  xlab("Type of FSA Crop") +
  ylab("Count")+
  theme_minimal()+
  theme(legend.position = "none")

ap_plot2
```

```{r}
# Make data ready for sankey 
coarse2 <- coarse %>% 
  group_by(CDL_Code) %>% 
  mutate(CDL_n = n()) %>% 
  group_by(FSA_Crop_Abbr) %>% 
  mutate(FSA_n = n()) %>% 
  group_by(FSA_Type_Name) %>%
  mutate(FSAType_n = n())

coarse2$CDL_Crop_Name[is.na(coarse2$CDL_Crop_Name)] <- "No Category"

class(coarse2$CDL_Code)

 coarse3 <- coarse2 %>% 
  filter(!CDL_Crop_Name %in% c("No Category", "Grapes", "Aquaculture", "Other Crops", "Herbs", "Misc. Vegs & Fruit")) %>%
  filter(!CDL_Code %in% c(60,237,254,241,230,233,231,232,226,235,234,239,240,228,225,238,236,26)) 
  #Talk to Andrea about removing other crops 


coarse3$FSA_Crop_Name <- factor(coarse3$FSA_Crop_Name, levels=unique(coarse3$FSA_Crop_Name))


sankey <- ggplot(coarse3, 
  aes(y = FSA_n, axis1 = reorder(FSA_Crop_Name, -FSA_n), axis2 = CDL_Crop_Name)) +
  geom_alluvium(aes(fill = CDL_Crop_Name), curve_type = "quintic")+
  geom_stratum(aes(fill = CDL_Crop_Name, color = "grey"), width = 1/8) + 
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_void()+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("FSA to CDL Coarsening")

#sankey <- sankey + 
  #ggrepel::geom_text_repel(aes(label = after_stat(stratum)),
        #stat = "stratum", size = 6, 
        #direction = "y", nudge_x = -.3)
#move labels with ggrepeltext: https://corybrunson.github.io/ggalluvial/articles/labels.html


sankey
```

```{r}
ggsave("./figs/sankey_all.png", width = 16, height = 25, units = "in")


svglite("./figs/sankey_all.svg", width = 8, height = 15)
sankey
dev.off()
```

## Let's try to filter based on number of FSA crop types to see better

```{r}
coarse_few <- coarse3 %>% filter(FSA_n < 20)
coarse_big <- coarse3 %>% filter(FSA_n > 20)

sankey_few <- ggplot(coarse_few, 
  aes(y = FSA_n, axis1 = reorder(FSA_Crop_Name, -FSA_n), axis2 = CDL_Crop_Name)) +
  geom_alluvium(aes(fill = CDL_Crop_Name), curve_type = "quintic")+
  geom_stratum(aes(fill = CDL_Crop_Name, color = "grey"), width = 1/8) + 
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_void()+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("FSA to CDL Coarsening")

#sankey <- sankey + 
  #ggrepel::geom_text_repel(aes(label = after_stat(stratum)),
        #stat = "stratum", size = 6, 
        #direction = "y", nudge_x = -.3)
#move labels with ggrepeltext: https://corybrunson.github.io/ggalluvial/articles/labels.html


sankey_few
ggsave("./figs/sankey_few.png", width = 16, height = 25, units = "in")
svglite("./figs/sankey_few.svg", width = 8, height = 15)
sankey_few
dev.off()


sankey_big <- ggplot(coarse_big, 
  aes(y = FSA_n, axis1 = reorder(FSA_Crop_Name, -FSA_n), axis2 = CDL_Crop_Name)) +
  geom_alluvium(aes(fill = CDL_Crop_Name), curve_type = "quintic")+
  geom_stratum(aes(fill = CDL_Crop_Name, color = "grey"), width = 1/8) + 
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_void()+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("FSA to CDL Coarsening")

#sankey <- sankey + 
  #ggrepel::geom_text_repel(aes(label = after_stat(stratum)),
        #stat = "stratum", size = 6, 
        #direction = "y", nudge_x = -.3)
#move labels with ggrepeltext: https://corybrunson.github.io/ggalluvial/articles/labels.html


sankey_big

ggsave("./figs/sankey_big.png", width = 16, height = 25, units = "in")
svglite("./figs/sankey_big.svg", width = 8, height = 15)
sankey_few
dev.off()
```

## Corn

```{r}
corn <- coarse2 %>% 
  filter(FSA_Crop_Abbr == "CORN")

corn_sankey <- ggplot(corn,
       aes(y = FSA_n, axis1 = FSA_Type_Name, axis2 = FSA_Crop_Name, axis3 = CDL_Crop_Name)) +
  scale_x_discrete(limits = c("FSA Type", "FSA Crop", "CDL Crop"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = CDL_Crop_Name), curve_type = "quintic")+
  geom_stratum()+ 
  geom_text(stat = "stratum", aes(label = after_stat(stratum), size = 6)) +
  theme_minimal()+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        #axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.x=element_blank())+
  #scale_fill_viridis_d() +
  ggtitle("FSA to CDL Coarsening - Corn")

corn_sankey
ggsave("./figs/sankeycorn.png", width = 13, height = 15, units = "in")

svglite("./figs/sankeycorn_svg.svg", width = 13, height = 15)
corn_sankey
dev.off()
```

## Herbs

```{r}
library(ggrepel)
herbs <- coarse2 %>% 
  filter(CDL_Code == 57)

herb_sankey <- ggplot(herbs,
       aes(y = FSA_n, axis1 = FSA_Crop_Name, axis2 = FSA_Type_Name, axis3 = CDL_Crop_Name)) +
  scale_x_discrete(limits = c("FSA Crop", "FSA Type", "CDL Crop"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = FSA_Type_Name), color="black", curve_type = "quintic", width = 1/20)+
  geom_stratum()+ 
  theme_minimal()+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        #axis.text.x=element_blank() - change size!
        axis.text.y=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("FSA to CDL Coarsening - Herbs")

herb_sankey <- herb_sankey + 
  ggrepel::geom_text_repel(aes(label = after_stat(stratum)),
        stat = "stratum", size = 6, 
        direction = "y", nudge_x = -.3)

herb_sankey
ggsave("./figs/sankeyherb.png", width = 13, height = 15, units = "in")
```

## Misc. Fruits and Veg

```{r}
fv_misc <- coarse2 %>% 
  filter(CDL_Code == 47)

miscfv_sankey <- ggplot(fv_misc,
       aes(y = FSA_n, axis1 = FSA_Crop_Name, axis2 = CDL_Crop_Name)) +
  scale_x_discrete(limits = c("FSA Crop", "CDL Crop"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = FSA_Crop_Name),aes.bind=TRUE, color="black", curve_type = "quintic", width = 1/10)+
  geom_stratum()+ 
  #geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal()+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        #axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("FSA to CDL Coarsening - Misc. Vegs & Fruit")


miscfv_sankey <- miscfv_sankey + 
  ggrepel::geom_text_repel(aes(label = after_stat(stratum)),
        stat = "stratum", size = 4, 
        direction = "y", nudge_x = -.3)

miscfv_sankey

ggsave("./figs/sankeyfvmisc.png", width = 15, height = 12, units = "in")
```

## Greens

```{r}
greens <- coarse2 %>% 
  filter(CDL_Code == 219)

greens_sankey <- ggplot(greens,
       aes(y = FSA_n, axis1 = FSA_Type_Name, axis2 = FSA_Crop_Name, axis3 = CDL_Crop_Name)) +
  scale_x_discrete(limits = c("FSA Type", "FSA Crop", "CDL Crop"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = FSA_Type_Name), aes.bind=TRUE, color="black", curve_type = "arctangent", width = 1/20)+
  geom_stratum()+ 
  theme_minimal()+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        #axis.text.x=element_blank() - change size!
        axis.text.y=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("FSA to CDL Coarsening - Greens")

greens_sankey <- greens_sankey + 
  ggrepel::geom_text_repel(aes(label = after_stat(stratum)),
        stat = "stratum", size = 6, 
        direction = "y", nudge_x = -.3)

greens_sankey
ggsave("./figs/sankeygreens.png", width = 13, height = 15, units = "in")


svglite("./figs/sankeygreens_svg.svg", width = 13, height = 15)
greens_sankey
dev.off()
```

## Other Crops

```{r}
other_crops <- coarse2 %>% 
  filter(CDL_Code == 44) %>% filter(!FSA_Crop_Code == 7501)

othercrops_sankey <- ggplot(other_crops,
       aes(y = FSA_n, axis1 = reorder(FSA_Crop_Name, -FSA_n), axis2 = CDL_Crop_Name)) +
  scale_x_discrete(limits = c("FSA_Crop", "CDL Crop"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = FSA_Crop_Name), color="black", curve_type = "arctangent", width = 1/20)+
  geom_stratum()+ 
  theme_minimal()+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        #axis.text.x=element_blank() - change size!
        axis.text.y=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("FSA to CDL Coarsening - Other Crops")

othercrops_sankey <- othercrops_sankey + 
  ggrepel::geom_text_repel(aes(label = after_stat(stratum)),
        stat = "stratum", size = 6, 
        direction = "y", nudge_x = -.3)

othercrops_sankey
ggsave("./figs/othercrops_sankey.png", width = 13, height = 21, units = "in")

svglite("./figs/sankeyothercrops_svg.svg", width = 13, height = 21)
othercrops_sankey
dev.off()

#Flowers
flowers <- coarse2 %>% 
  filter(CDL_Code == 44) %>% filter(FSA_Crop_Code == 7501)

flowers_sankey <- ggplot(flowers,
       aes(y = FSA_n, axis1 = FSA_Type_Name, axis2=FSA_Crop_Name, axis3 = CDL_Crop_Name)) +
  scale_x_discrete(limits = c("FSA Type", "FSA Crop", "CDL Crop"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = FSA_Type_Name), color="black", curve_type = "arctangent", width = 1/20)+
  geom_stratum()+ 
  theme_minimal()+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        #axis.text.x=element_blank() - change size!
        axis.text.y=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("FSA to CDL Coarsening - Other Crops")

flowers_sankey <- flowers_sankey + 
  ggrepel::geom_text_repel(aes(label = after_stat(stratum)),
        stat = "stratum", size = 6, 
        direction = "y", nudge_x = -.3)

flowers_sankey
ggsave("./figs/flowers_sankey.png", width = 13, height = 15, units = "in")


svglite("./figs/sankeyflowers_svg.svg", width = 13, height = 15)
flowers_sankey
dev.off()

##COME BACK TO
```

## Other tree crops

```{r}
other_tree <- coarse2 %>% 
  filter(CDL_Code == 71)

othertree_sankey <- ggplot(other_tree,
       aes(y = FSA_n, axis1 = FSA_Crop_Name, axis2 = CDL_Crop_Name)) +
  scale_x_discrete(limits = c("FSA Crop", "CDL Crop"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = FSA_Crop_Name), color="black", curve_type = "quintic", width = 1/20)+
  geom_stratum()+ 
  theme_minimal()+
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        #axis.text.x=element_blank() - change size!
        axis.text.y=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("FSA to CDL Coarsening - Other Tree Crops")

othertree_sankey <- othertree_sankey + 
  ggrepel::geom_text_repel(aes(label = after_stat(stratum)),
        stat = "stratum", size = 6, 
        direction = "y", nudge_x = -.3)

othertree_sankey
ggsave("./figs/othertree_sankey.png", width = 13, height = 20, units = "in")


svglite("./figs/sankeyothertree_svg.svg", width = 13, height = 20)
othertree_sankey
dev.off()
```
